# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data
```{r}
data<-read.csv("activity.csv")
#data$steps<-as.numeric(data$steps)
summary(data)
```


## What is mean total number of steps taken per day?
### Calculating the media by day and plotting the histogram
```{r fig.width=10}
data2<-na.omit(data)
mean_each_day<-tapply(data2$steps, data2$date, mean)

hist(mean_each_day,main="Mean Steps by Day",xlab="Mean Each Day")
```
### Calculating the global media and median
```{r}
mean(mean_each_day,na.rm=TRUE)
median(mean_each_day,na.rm=TRUE)

```



## What is the average daily activity pattern?
### Aggregating data by intervals and plotting
```{r fig.width=10}
mean_interval<-aggregate(data2$steps, by=list(data2$interval), 
                    FUN=mean, na.rm=TRUE)
colnames(mean_interval)<-c("interval","steps")
plot(mean_interval$interval,mean_interval$steps,type="l",xlab="5-minute interval",ylab="Average All Days",main="Average Daily Activity Pattern")
```
### Getting and showing the max of steps in the media interval dataset
```{r}
max_interval<-mean_interval[which.max( mean_interval$steps ),]
colnames(max_interval)<-c("Interval","Mean Steps")
max_interval

```
## Imputing missing values
### Number of missing values
```{r}
sum(!complete.cases(data))
```
### Imputing missing values by replacing them with the the mean of the 5-minute interval
### Steps
- Get a case complete dataset
- get the NA data set and replace the null values with the average of the given interval
- Put the imputed dataset with the complete dataset
- Show the characteristics of the new dataset (No NA's found)
```{r}
data_complete<-na.omit(data)
data_incomplete<-data[is.na(data$steps),2:3]
data_incomplete<- merge(data_incomplete,mean_interval,by="interval")
data_complete <- rbind(data_complete, data_incomplete)
str(data_complete)
sum(!complete.cases(data_complete))
```
### Basic statistics and Histogram for the new dataset

```{r fig.width=10}
mean_each_day_complete<-tapply(data_complete$steps, data_complete$date, mean)

hist(mean_each_day_complete,main="Mean Steps by Day",xlab="Mean Each Day")

mean(mean_each_day_complete)
median(mean_each_day_complete)
```

## Are there differences in activity patterns between weekdays and weekends?
### Creating the new var day_of_week 
```{r}
Sys.setlocale("LC_TIME", "English")
data_complete$day_of_week<-weekdays(as.Date(data_complete[,2]))
data_complete$day_of_week[which(data_complete$day_of_week == "Sunday" | data_complete$day_of_week == "Saturday")]<-"weekend"
data_complete$day_of_week[which(data_complete$day_of_week!="weekend")]<-"weekday"
data_complete$day_of_week<-as.factor(data_complete$day_of_week)

summary(data_complete)
```
## aggregating and Plotting the data

```{r fig.width=10}
mean_interval<-aggregate(data_complete$steps, by=list(data_complete$interval,data_complete$day_of_week),FUN=mean, na.rm=TRUE)

colnames(mean_interval)<-c("interval","day_of_week","steps")

library(lattice)
xyplot(mean_interval$steps~as.numeric(mean_interval$interval)|mean_interval$day_of_week, layout = c(1, 2),type="l",xlab="Interval",ylab="Number of Steps") 


```
